---
title: "Exploratory data analysis"
author: "Yohan Min"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    number_sections: yes
    self_contained: yes
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document:
    number_sections: yes
    toc: yes
    fig_cap: yes
    keep_tex: yes
  html_document:
    toc: yes
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}
# Install pacman

if (!require("pacman")) install.packages("pacman")
# p_load function loads packages if installed, or install then loads otherwise
pacman::p_load(dplyr, knitr, kableExtra, readr, pander, ggplot2,tidyr, mlmRev, stargazer,mgcv, 
               lme4,sjPlot, sf, purrr,broom.mixed,MASS,stringr)

data <- read_csv("turnover_babushkin.csv")


```

## Model in mathematical notation

Using the Babushkin Employee Turnover data, the brief research paper tried to perform an exploratory data analysis, and model `tenure` with respect to covariates, `age`, and `extraversion` with a group effect, `profession`. 


$$
Y_i\sim Normal(\mu_i,\sigma_{Y}^2),\\
\mu_i =\alpha_{j}+X_i\beta, \\
\alpha_j \sim Normal(\mu_{\alpha}, \sigma_{\alpha}^2).
$$

where $Y_i$ is a dependent variable, `tenure` for an observation i. It follows a normal distribution with a dispersion parameter $\sigma_{Y}$. $\alpha_{j}$ is a random intercept for a group j following a normal distribution with parameters, $\mu_{\alpha}$, and $\sigma_{\alpha}^2$. $β$ is covariate coefficient for $X_i$. 


## Exploratory plot

```{r}


data %>% 
  filter(profession %in% unique(data$profession)[-7]) %>% 
  ggplot(aes(x = age, y = tenure, color = factor(profession)))+
  geom_point()

```



<br>

## Estimating the model and reporting a table of parameter estimates

The method of estimation is a linear mixed-effects. 



```{r}

lm <- lmer(tenure ~ age+extraversion+ (1 |profession), data) # # random effect

tab_model(lm)


```

<br>

## Generalized Additive Model

Models impose assumptions to gain traction over data.Sometimes
the assumptions are close in spirit to the data and thus worth making to
simplify the patterns in the data. In other cases, even slightly weakening the assumptions
of a model strongly changes the results. To investigate the robustness of the results, I replaced the covariates with
“smooths” of the respective variables, and refit the model.


```{r}

gam <- gam(tenure ~ s(age)+s(extraversion)+
           s(profession, bs = "re"), data = data %>% 
             mutate(profession = factor(profession)), method = "REML") 

```




## Graphical representation 

I'm showing graphically whether using a smoother “makes a difference” substantively;
i.e., whether or not the response variable reacts to the smoothed variable in an
approximately linear fashion.I have reason to doubt the appropriateness or usefulness of my original model as one of the significant predictor, `age` shows non-linear pattern as below.

 

```{r}

plot(gam, all.terms = TRUE, page = 1)
plot(gam,  scheme = 1,select = 1,shade = TRUE, shade.col = "hotpink", 
     shift = coef(gam)[1])

```

In addition, the second model has lower `AIC` while higher `BIC` values than the original model. 

```{r}

IC <- rbind(list(lm, gam) %>% 
  lapply(AIC) %>% unlist(),
  list(lm, gam) %>% 
  lapply(BIC) %>% unlist())

rownames(IC) <- c("AIC", "BIC")
colnames(IC) <- c("GLM", "GAM")

IC %>% 
  round(2) %>% 
  kable() %>%
  kable_styling(full_width = FALSE, position = "left", 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


<br>

## Transformation 

Since I doubted the original model’s treatment of explanatory variable, I tried 
a simple transformation, that approximates
the smooth function produced by GAM. 

<br>

(1) refitting the model

```{r}


lm1 <- lmer(tenure ~ age+poly(extraversion,3) + (1 |profession), data) # # random effect

tab_model(lm1)

```

<br>

(2) reformulated  model fits better than the original model

```{r}


IC <- rbind(list(lm, lm1) %>% 
  lapply(AIC) %>% unlist(),
  list(lm, lm1) %>% 
  lapply(BIC) %>% unlist())


rownames(IC) <- c("AIC", "BIC")
colnames(IC) <- c("Original", "Modified")

IC %>% 
  round(2) %>% 
  kable() %>%
  kable_styling(full_width = FALSE, position = "left", 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```


<br> 

(3) the similarity of the new fit to the smoothed model

```{r}

# original model 
x <- 
  data %>%                  
  dplyr::select(age,extraversion) %>% 
  summarize_all(mean) %>%   # summarize the means for most variables (can't average Species)
  uncount(20) %>% # repeat the row 10 times
  mutate(age = seq(min(data$age), max(data$age), length.out = 20),
         const = 1) %>% 
  dplyr::select(const, age, extraversion) %>% 
  as.matrix()


sims <- 10000
pe <- lm %>% tidy() %>% dplyr::select(estimate) %>% as.matrix()
pe <- pe[1:3]
vc <- vcov(lm)
simbetas <- mvrnorm(sims, pe, vc)

xbeta <- x %*% t(simbetas)


pe <- apply(xbeta, 1, mean) 
upper <- apply(xbeta, 1, quantile, probs= 0.975)
lower <- apply(xbeta, 1, quantile, probs= 0.025) 

or_pe <- cbind(x[,2],pe,upper,lower) %>%
  as.data.frame() %>% 
  mutate(model = "orginal")


# Gam model

new <- x %>% 
  as.data.frame() %>% 
  dplyr::select(-const) %>% 
  mutate(profession = "HR")

pr <- predict(gam, se.fit = T, type = "response", newdata = new) 
gam_pe <- data.frame(age = x[ ,2], pe = pr$fit, upper = pr$fit + 1.96*pr$se.fit,
           lower = pr$fit - 1.96*pr$se.fit,
           model = "GAM")


library(bootpredictlme4)
preds <- predict(lm1, newdata= new, re.form=NA, se.fit=TRUE, nsim=50)
md_pe <- data.frame(age = x[ ,2], pe = preds$fit, upper = preds$fit + 1.96*preds$se.fit,
           lower = preds$fit - 1.96*preds$se.fit,
           model = "modified")


rbind(or_pe %>% 
          rename(age = V1), md_pe,gam_pe) %>%
  ggplot() +
  aes(y = pe,
      x = age,
      ymax = upper,
      ymin = lower,
      color = model) +
  geom_line(size = 0.7)+
  geom_ribbon(aes(fill = model), alpha = 0.2) +
  guides(color="none") + 
  labs(y = "tenure", fill = "Model") +
  theme_minimal() 

```



