---
title: "R Markdown"
author: '[Yohan Min](https://reconjohn.github.io/)'
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
    number_sections: true
    code_folding: hide
    self_contained: yes
  pdf_document:
    number_sections: yes
    toc: yes
    fig_cap: yes
    keep_tex: yes
---

<!--these following settings control width and also make a DIV tag for black outlines for figures-->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# load several packages
pacman::p_load(readr, tidyverse, magrittr, knitr, kableExtra, readstata13, stargazer, pander, captioner, keyring, HMDHFDplus, flextable)

# for captions and cross references
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```


<!-- for a wide presentation -->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<!-- a DIV for black outline, good for figures-->
<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>

# Problem 
Using the Babushkin employment data, create a ggplot graphic (`r figure_nums(name = "boxplot", display = "cite")`) that shows the relationship between the variables gender and anxiety.

```{r}

# download data
df <- read_csv(file = "https://raw.githubusercontent.com/teuschb/hr_data/master/datasets/turnover_babushkin.csv")

# anxiety by gender 
df %>% 
  mutate(gender = factor(gender, levels = c("m", "f"))) %>%
  ggplot(aes(x = gender, y = anxiety)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=1, size=5, color="red", fill="red")

```

_`r figure_nums(name = "boxplot", caption = "Relationship between gender and anxiety")`_

Comment: There is a difference in anxiety between gender. 
\

# Problem 
Using the Babushkin employment data, create a table that includes the count and percent of records with the combination of the variables selfcontrol and independ, arranged in descending order of count (`r table_nums(name = "table", display = "cite")`).

```{r}

df %>% 
  group_by(selfcontrol, independ) %>% 
  summarise(n = n(), .groups = "drop") %>% # degroup if necessary 
  mutate("%" = round(n / sum(n) * 100, 2)) %>% 
  arrange(desc(n)) %>% 
  pander()
```

_`r table_nums(name = "table", caption = "Count and percent of variabels, selfcontrol and independ")`_

Comment: The table shows negatively weak correlation between the variables, `selfcontrol` and `independ`. 

# Problem 
Using the Babushkin employment data, create a ggplot graphic that shows the relationship between the variables selfcontrol and independ (`r figure_nums(name = "ggplot", display = "cite")`). These are rounded to the 10ths place, therefore there are a number of records that have the same combination of values for the two variables. Because of this, a standard scatter plot will result in overprinting of points that have the same value for the two variables. Try to find a method that presents a better view of the distribution of the data.

```{r}
df %>% 
  ggplot(aes(x = selfcontrol, y = independ))+
  geom_point(alpha=.5, position=position_jitter(h=.3, w=.3))+
  geom_smooth()

```

_`r figure_nums(name = "ggplot", caption = "Relationship between the variables selfcontrol and independ")`_

Comment:  The figure confirms the neagtively weak correlation. 


# Problem 
Using the Human Mortality data, create a ggplot line graph (`r figure_nums(name = "gg1", display = "cite")`, and `r figure_nums(name = "gg2", display = "cite")`) that includes a time series of mortality data based on the 1 x 1 death rates for three countries. Explain any patterns you see based on what you know about the characteristics and histories of these three countries.

```{r}

# load required packages
library(HMDHFDplus)
library(keyring)
library(tidyverse)

# Set your password for Human Mortality Database (HMD)
# does a key exist? if not, create one
if (key_list(service = "human-mortality-database") %>% nrow() == 0) {
  keyring::key_set(
    service = "human-mortality-database",
    username = "min25@uw.edu"
  )
  # Enter your HMD password in the prompt
}


# Function to download a specified HMD dataset item for a single county
read_hmd_country <- function(CNTRY, item) {
  HMDHFDplus::readHMDweb(
    CNTRY = CNTRY,
    item = item,
    username = keyring::key_list("human-mortality-database")$username,
    password = keyring::key_get(
      service = "human-mortality-database",
      username = keyring::key_list("human-mortality-database")$username
    )
  )
}

# Help function to list the available countries
countries <- HMDHFDplus::getHMDcountries()

# Download a dataset iteratively for all countries using purrr::map()
# In this case, age-specific mortality in 1-year periods x 1-year age groups
# for all 1-year periods available
mx_1x1 <- countries %>%
  # Returns a list of data.frames, adding a column for country code to each
  purrr::map(function(country) {
    read_hmd_country(country, "Mx_1x1") %>%
      dplyr::mutate(country = country)
  }) %>%
  # Combines the data.frames into a single data.frame
  dplyr::bind_rows()


mx_1x1 %>% 
  filter(country %in% c("KOR","JPN","USA")) %>% 
  filter(Year > 2012 & Year < 2019) %>% 
  gather("gender", "value", Female, Male) %>% 
  ggplot(aes(x = Age, y = value, group = Year, color = gender)) +
  geom_line(size = 1)+
  facet_grid(Year~ country) + 
  labs(y = "Mortality from 2013 - 2018")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 20)) + 
  theme(axis.title=element_text(size=15),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        strip.text = element_text(size = 13))
```


_`r figure_nums(name = "gg1", caption = "Mortality in Japan, Korea, and USA from 2013 to 2018")`_

Comment: Korea seems to have higher mortality at order age groups than Japan and USA. 

```{r}

mx_1x1 %>% 
  filter(country %in% c("KOR","JPN","USA")) %>% 
  filter(Year > 2012 & Year < 2019) %>% 
  gather("gender", "value", Female, Male) %>% 
  group_by(country, gender, Age) %>% 
  summarise(mean = mean(value)) %>% 
  ggplot(aes(x = Age, y = mean, group = country, color = country)) +
  geom_line(size = 1)+
  scale_x_continuous(breaks = seq(0, 115, by =10),)+
  facet_wrap(~ gender)+
  labs(y = "Average Mortality from 2013 - 2018")+
  theme_bw()
```


_`r figure_nums(name = "gg2", caption = "Mortality in Japan, Korea, and USA from 2013 to 2018")`_

Comment: Among the three countries, Japan presents the lowest mortality for both genders from years 2013 to 2018. 

# Source code
```{r comment=''}
cat(readLines("csde502_2022_assignment02_Yohan.Rmd"), sep = '\n')
```