---
title: "Functions and sampling "
author: '[Yohan Min](https://reconjohn.github.io/), min25@uw.edu'
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 6
    number_sections: yes
    code_folding: hide
    self_contained: yes
always_allow_html: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = F)

options(tigris_use_cache = TRUE)
pacman::p_load(demogR, demography, magrittr, knitr, kableExtra, readstata13, captioner,tigris, sf, tidyverse)

table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```

Regarding the `r figure_nums(name = "hist", display = "cite")`, all the answers are as follows. 


```{r}

gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),
rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))

barplot(table(gradpain), las=1, ylab="Frequency", xlab="Pain Scale",
main="How painful is graduate school?")


```

_`r figure_nums(name = "hist", caption = "Survey results on pain of graduate school")`_

<br> 

# How many graduate students are in the sample?

```{r}

# length(gradpain)
s_num <- table(gradpain) %>% sum()

```
There are `r s_num` students in the sample. 

<br>

# What is the sample mean?

```{r}

s_mean <- mean(gradpain)


```

The mean is `r s_mean`. 

```{r}


s_fun <- function(d.vec, n){
  result <- sample(x = d.vec, size = n, replace = TRUE)
  list("size" = n, "mean" = mean(result))
}
  
set.seed(7)

fin <- c()
for(i in 1:100){
  result <- s_fun(gradpain, s_num)
  fin <- c(fin, result$mean)
}

  
```

# What is the mean of these 100 means?

The mean of the 100 means is `r round(mean(fin),2)`. 

<br> 

# What is the standard deviation of these 100 means?

The standard deviation of these 100 means is `r round(sd(fin),2)`.

```{r}

ss_fun <- function(d.vec, n, num.samples){
  fin <- c()
  for(i in 1:num.samples){
    result <- sample(x = d.vec, size = n, replace = TRUE)
    fin <- c(fin, mean(result))
  }
  
      list("size" = n, "total size" = n*num.samples, "means" = fin,  "mean" = mean(fin), "sd" = sd(fin), 
         "95% CI" = quantile(fin, probs = c(0.25, 0.975)))
}


set.seed(7)

ans <- ss_fun(d.vec = gradpain, n = length(gradpain), num.samples = 100)
  

```


# What does your function return for the mean of means?

The mean of means is `r round(ans$mean,2)`. 

<br>

# What does your function return for the standard deviation of means?

The standard deviation of means is `r round(ans$sd,2)`.

<br>

# What does your function return for the 95% confidence interval around the mean?

The 95% confidence interval around the mean is `r round(ans$'95% CI',2)` respectively. 

<br>

# Source code
```{r comment=''}
cat(readLines("csde502_2022_assignment04_Yohan.Rmd"), sep = '\n')
```


