---
title: "Introduction to Add Health"
author: 'Yohan Min, min25@uw.edu'
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 6
    number_sections: yes
    self_contained: yes
always_allow_html: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = F)

options(tigris_use_cache = TRUE)
pacman::p_load(demogR, demography, magrittr, knitr, kableExtra, readstata13, captioner,tigris, sf, tidyverse)

table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```



## Generate a new data frame from the full public Add Health data set (http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip) that consists of a subset of at least four of the original columns. Do not use any of the variables used as examples in Lesson 7.


```{r}


# download and unzip the larger data set
myUrl <- "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip"
mytempdir <- tempdir()

# zip file in $temp -- basename gets just the file name from the URL and not the URL path;
#   file.path stitches the tempdir() path to the file name
zipfile <- file.path(mytempdir, basename(myUrl))

# dta file in $temp
dtafile <- tools::file_path_sans_ext(zipfile)

# check if the dta file exists
if (!file.exists(dtafile)) {
    # if the dta file doesn't exist, check for the zip file
    # check if the zip file exists, download if necessary
    if (!file.exists(zipfile)) {
        curl::curl_download(url = myUrl, destfile = zipfile)
    }
    # unzip the downloaded zip file
    if (file.exists(zipfile)) {
        unzip(zipfile = zipfile, exdir = mytempdir)
    }
}

# if the data set has not been read, read it in
if (!exists("ahcomplete")) {
    data <- haven::read_dta(dtafile)
}
# lowercase column names
colnames(data) %<>% str_to_lower()

data <- data[,c("h1gi15","h1gi1m", "h1gi1y", "h1gi16y")]

# data %>%
#   attributes()
# 
# data %>% purrr::map(function(x) attributes(x))


new_data <- data.frame("married" = factor(data$h1gi15, labels = data$h1gi15 %>% 
                                                 attributes() %>% extract2("labels") %>%
                                                 names(),ordered = TRUE),
                       "birth_year" = factor(data$h1gi1y, labels = data$h1gi1y %>% 
                                               attributes() %>% extract2("labels") %>%
                                               names(),ordered = TRUE),
                       "birth_month" = factor(data$h1gi1m, labels = data$h1gi1m %>% 
                                                attributes() %>% extract2("labels") %>%
                                                names(),ordered = TRUE),
                       "married_year" = factor(data$h1gi16y, labels = data$h1gi16y %>% 
                                                 attributes() %>% extract2("labels") %>%
                                                 names(),ordered = TRUE))

```


## The code should save the data frame as an RDS file in a temporary folder, which can be specified in R as tempdir().


```{r}

saveRDS(object = new_data, file = file.path(mytempdir, "data.RDS"))

```



## Create some frequency tables:

### Create a frequency table from each variable, using both counts and percentages.

_`r table_nums(name = "fr", caption = "Married frequency tables of counts and percentages")`_

```{r}

  
  new_data %>%
    group_by(married) %>%
    summarise(n = n(), .groups = "drop") %>% 
    mutate(percent = n/ sum(n)) %>%
    kable() %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

_`r table_nums(name = "fr3", caption = "Birth year frequency tables of counts and percentages")`_

```{r}
  
  
    new_data %>%
    group_by(birth_year) %>%
    summarise(n = n(), .groups = "drop") %>% 
    mutate(percent = n/ sum(n)) %>%
    kable() %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

_`r table_nums(name = "fr4", caption = "Birth month frequency tables of counts and percentages")`_

```{r}
    
    
      new_data %>%
    group_by(birth_month) %>%
    summarise(n = n(), .groups = "drop") %>% 
    mutate(percent = n/ sum(n)) %>%
    kable() %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

_`r table_nums(name = "fr5", caption = "Married year frequency tables of counts and percentages")`_

```{r}
      
      
        new_data %>%
    group_by(married_year) %>%
    summarise(n = n(), .groups = "drop") %>% 
    mutate(percent = n/ sum(n)) %>%
    kable() %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  



```



### Create a frequency table based on at least two variables, also with counts and percentages.

_`r table_nums(name = "fr1", caption = "Two Variables frequency tables of counts and percentages")`_

```{r}

tb <- new_data %>%
  group_by(birth_year, married_year) %>%
  summarise(n = n(), .groups = "drop") %>% 
  mutate(percent = n/ sum(n)) 


tb %>%
  kable() %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```



## Create at least one graph from data in the data frame.


```{r}


ggplot(data = tb, aes(x = birth_year, y = n, fill = married_year)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(~married_year) +
  labs(x = "Birth year", y = "Number of respondents")


```


_`r figure_nums(name = "hist", caption = "Number of respondents by birth year and marred year")`_

<br> 



# Source code
```{r comment=''}
cat(readLines("csde502_2022_assignment07_Yohan.Rmd"), sep = '\n')
```


