---
title: "File naming example"
author: "[CSDE 502](https://csde-uw.github.io/csde502-winter-2022/)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output:
    html_document:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue
---

```{r setup, include=FALSE}
pacman::p_load(knitr, kableExtra, tidyverse, magrittr)
knitr::opts_chunk$set(echo = FALSE)
```

The following code chunk creates a set of dummy files in a temporary folder, pausing for one second after each file is created (to affect the file modification time--so running this code will take longer than expected). It will open a file explorer on Windows and Mac computers and also display an interactive table allowing sorting by file name.

```{r}
# make a dummy directory
dud_dir <- normalizePath(file.path(tempdir(), "delete_me_later"), winslash = "/")

if(!dir.exists(dud_dir)){
    dir.create(path = dud_dir)
}

# make a dummy a data frame to write to a file
x <- data.frame(foo = "bar")

# make some stupidly named files, Dec 16, 2021
write.csv(x = x, file = file.path(dud_dir, "file 121621.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 12_16_21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 12-16-21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 16-12-21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_121621.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)

# make some stupidly named files, Dec 17, 2021
write.csv(x = x, file = file.path(dud_dir, "file 121721.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 12_17_21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 12-17-21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file 17-12-21.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_121721.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)

# write two logically named files
write.csv(x = x, file = file.path(dud_dir, "file_20211217.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_20211216.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_2021-12-17.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_2021-12-16.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_2021_12_17.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)
write.csv(x = x, file = file.path(dud_dir, "file_2021_12_16.csv"), quote = FALSE, row.names = FALSE); Sys.sleep(1)

# get the OS
get_os <- function(){
  sysinf <- Sys.info()
  if (!is.null(sysinf)){
    os <- sysinf['sysname']
    if (os == 'Darwin')
      os <- "osx"
  } else { ## mystery machine
    os <- .Platform$OS.type
    if (grepl("^darwin", R.version$os))
      os <- "osx"
    if (grepl("linux-gnu", R.version$os))
      os <- "linux"
  }
  tolower(os)
}
os <- get_os()

# open a file explorer
if(os == "windows"){
    shell.exec(dud_dir)
}
if(os == "osx"){
    system2("open", dud_dir)
}

# get the list of files
fnames <- tibble(
    path = list.files(path = dud_dir,
                     pattern = "*.csv", 
                     full.names = TRUE)
    )
fnames %<>% 
        mutate(fname = path %>% basename(),
               mod_time = map_dfr(path, file.info)$mtime %>% as.character())

# get the file creation dates


# print with DT for sorting
DT::datatable(fnames, options = list(pageLength = nrow(fnames)))
```

Which file is newer?

You should be convinced that yyyymmdd is a better way to name files by date. Your reward will be to delete `r dud_dir`.